Relevance-based Quantization of Scattering Features for Unsupervised Mining of Environmental Audio

### 2018-01-18 paper/structureScene.tex
cleaning

\subsection{Gammatone wavelets}
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{bw/gammatones}
\caption{
\label{fig:gammatones}
Gammatone wavelets $\psi(t)$ in the time domain with quality factors (a) $Q = 4$ and (b) $Q = 1$.
Oscillations represent the real and imaginary parts. The envelope represents the complex modulus.}
\end{center}
\end{figure}
Wavelets
$\boldsymbol{\psi_{\gamma_1}}(t)$ and $\boldsymbol{\psi_{\gamma_2}}(t)$ are designed as fourth-order Gammatone
wavelets with one vanishing moment \cite{Venkitaraman2014}, and are shown in Figure \ref{fig:gammatones}.
In the context of auditory scene analysis, the asymmetric envelope of Gammatone wavelets is more biologically plausible than the symmetric, Gaussian-like envelope of the more widely used Morlet wavelets.
Indeed, it allows to reproduce two important psychoacoustic effects in the mammalian cochlea: the asymmetry of temporal masking and the asymmetry of spectral masking  \cite{Fastl2007}.
The asymmetry of temporal masking is the fact that a masking noise has to be louder if placed after the onset of a stimulus rather than before.
Likewise, because critical bands are skewed towards higher frequencies, a masker tone has to be louder if it is above the stimulus in frequency rather than below.
It should also be noted that Gammatone wavelets follow the typical amplitude profile of natural sounds, beginning with a relatively sharp attack and ending with a slower decay.
As such, they are similar to filters discovered automatically by unsupervised encoding of natural sounds \cite{Smith2006}.
This suggests that, despite being hand-crafted and not learned, Gammatone wavelets provide a sparser time-frequency representation of acoustic scenes compared to other variants.

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{bw/compression}
\caption{
\label{fig:histograms}
Histogram of values taken by the first-order scattering coefficient $\mathbf{S}\boldsymbol{x}(\gamma)$, corresponding to a center acoustic frequency of $302\,\mathrm{Hz}$,
(a) before and (b) after logarithmic compression.}
\label{fig:compression}
\end{center}
\end{figure}

@@

The implementations of the presented methods and the experimental protocol are available online.\footnote{\url{https://github.com/mathieulagrange/tetci17}} % \ja{What is ``the one''?}

@@

%For most of the experiments reported in this paper, we will consider the DCASE2013 dataset \cite{giannoulis2013database, 7100934}. Larger datasets exists: the
%Rouen dataset \cite{rakotomamonjy2015histogram} and the public part of the DCASE2016 challenge.%
The experiments in this paper are carried out on the DCASE 2013 \cite{7100934} and DCASE 2016 \cite{Mesaros2016_EUSIPCO} datasets.
The DCASE 2013 dataset consists of two parts, namely a public and a private subset, each made up of $100$ $30$-second recordings of various acoustic scenes, sampled at $44100\,\mathrm{Hz}$. The $100$ recordings are evenly divided into $10$ acoustic scene classes. To build the DCASE 2013 dataset, three different recordists visited a wide variety of locations in Greater London over a period of several months and in each scene recorded a few minutes of audio. No
systematic variations in the recordings covaried with scene
type: all recordings were made under moderate weather conditions, at varying times of day and week, and each recordist recorded each scene type. As a consequence, DCASE 2013 dataset enjoys an interesting intra-class diversity while remaining manageable in terms of size, making it suitable for extensive evaluation of algorithmic design choices \cite{lagrange:hal-01082501}.
In addition, it is still a challenging dataset, with a state-of-the-art system based using handcrafted features as input to SVMs achieving $76\%$ \cite{roma2013} (winner of the DCASE2013 challenge), while recent approaches based on label tree embedding achieve between $84\%$ and $87\%$ on DCASE 2013 depending on the prior knowledge used during training \cite{phan2016label}.

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{bw/unsupervised_test2}
\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for MFCCs and scattering with logarithmic compression, as a function of the rank $k$.}
\label{fig:ASS_1}
\end{center}
\end{figure}

@@

%\bibliographystyle{spmpsci}
%\bibliography{biblio}
%\printbibliography

>>>>

\subsection{Gammatone wavelets}
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{gfx/gammatones}
\caption{
\label{fig:gammatones}
Gammatone wavelets $\psi(t)$ in the time domain with quality factors (a) $Q = 4$ and (b) $Q = 1$.
Oscillations represent the real and imaginary parts. The envelope represents the complex modulus.}
\end{center}
\end{figure}
Wavelets
$\boldsymbol{\psi_{\gamma_1}}(t)$ and $\boldsymbol{\psi_{\gamma_2}}(t)$ are designed as fourth-order Gammatone
wavelets with one vanishing moment \cite{Venkitaraman2014}, and are shown in Figure \ref{fig:gammatones}.
In the context of auditory scene analysis, the asymmetric envelope of Gammatone wavelets is more biologically plausible than the symmetric, Gaussian-like envelope of the more widely used Morlet wavelets.
Indeed, it allows to reproduce two important psychoacoustic effects in the mammalian cochlea: the asymmetry of temporal masking and the asymmetry of spectral masking  \cite{Fastl2007}.
The asymmetry of temporal masking is the fact that a masking noise has to be louder if placed after the onset of a stimulus rather than before.
Likewise, because critical bands are skewed towards higher frequencies, a masker tone has to be louder if it is above the stimulus in frequency rather than below.
It should also be noted that Gammatone wavelets follow the typical amplitude profile of natural sounds, beginning with a relatively sharp attack and ending with a slower decay.
As such, they are similar to filters discovered automatically by unsupervised encoding of natural sounds \cite{Smith2006}.
This suggests that, despite being hand-crafted and not learned, Gammatone wavelets provide a sparser time-frequency representation of acoustic scenes compared to other variants.

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{gfx/compression}
\caption{
\label{fig:histograms}
Histogram of values taken by the first-order scattering coefficient $\mathbf{S}\boldsymbol{x}(\gamma)$, corresponding to a center acoustic frequency of $302\,\mathrm{Hz}$,
(a) before and (b) after logarithmic compression.}
\label{fig:compression}
\end{center}
\end{figure}

@@

The implementations of the presented methods and the experimental protocol are available online.\footnote{\url{https://github.com/mathieulagrange/paperRelevanceBasedSimilarity}} % \ja{What is ``the one''?}

@@

%For most of the experiments reported in this paper, we will consider the DCASE2013 dataset \cite{giannoulis2013database, 7100934}. Larger datasets exists: the
%Rouen dataset \cite{rakotomamonjy2015histogram} and the public part of the DCASE2016 challenge.%
The experiments in this paper are carried out on the DCASE 2013 \cite{7100934} dataset.
The DCASE 2013 dataset consists of two parts, namely a public and a private subset, each made up of $100$ $30$-second recordings of various acoustic scenes, sampled at $44100\,\mathrm{Hz}$. The $100$ recordings are evenly divided into $10$ acoustic scene classes. To build the DCASE 2013 dataset, three different recordists visited a wide variety of locations in Greater London over a period of several months and in each scene recorded a few minutes of audio. No
systematic variations in the recordings covaried with scene
type: all recordings were made under moderate weather conditions, at varying times of day and week, and each recordist recorded each scene type. As a consequence, DCASE 2013 dataset enjoys an interesting intra-class diversity while remaining manageable in terms of size, making it suitable for extensive evaluation of algorithmic design choices \cite{lagrange:hal-01082501}.
%In addition, it is still a challenging dataset, with a state-of-the-art system based using handcrafted features as input to SVMs achieving $76\%$ \cite{roma2013} (winner of the DCASE2013 challenge), while recent approaches based on label tree embedding achieve between $84\%$ and $87\%$ on DCASE 2013 depending on the prior knowledge used during training \cite{phan2016label}.

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=\columnwidth]{gfx/unsupervised_test2}
\caption{Acoustic scene similarity retrieval in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for MFCCs and scattering with logarithmic compression, as a function of the rank $k$.}
\label{fig:ASS_1}
\end{center}
\end{figure}

@@

%\bibliographystyle{spmpsci}
%\bibliography{biblio}
\printbibliography
